---
title: "BBC Women of 2020"
subtitle: "Visualizing women's achievement across continents with sunburstR"
author: "Paul Apivat Hanvongse"
date: "12/17/2020"
output: 
    html_document:
        toc: true
        toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries

The two main libraries are `tidyverse` (mostly `dplyr` so you can just load that if you want) and `sunburstR`. There are other packages for sunburst plots including: [plotly](https://plotly.com/r/sunburst-charts/) and [ggsunburst (of ggplot)](https://github.com/didacs/ggsunburst), but we'll explore [sunburstR](https://www.rdocumentation.org/packages/sunburstR/versions/2.1.5/topics/sunburst) in this post.

```{r libraries, message=FALSE}
library(tidyverse)
library(sunburstR)
```

## Load Data & Explore

The data is from [week 50 of TidyTuesday](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-12-08/readme.md), exploring the BBC's top 100 influential women of 2020. 


```{r exploratory, results='hide', message=FALSE}
women <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-12-08/women.csv')

head(women)
```

## Add Continents

The original dataset organized 100 women by category, country, role and description. I found that for employing the sunburst plot, I would want to group countries together by continents. 

I manually added country names to continent vectors, then added a new column to the `women` dataframe to conditionally add continent name.

We could then focus on six continents rather than 65 separate countries. 

```{r continents, results='hide'}
# add continent as character vector
asia <-  c('Afghanistan', 'Bangladesh', 'China', 'Exiled Uighur from Ghulja (in Chinese, Yining)', 'Hong Kong', 'India', 'Indonesia', 'Iran', 'Iraq/UK', 'Japan', 'Kyrgyzstan', 'Lebanon', 'Malaysia', 'Myanmar', 'Nepal', 'Pakistan', 'Singapore', 'South Korea', 'Syria', 'Thailand', 'UAE', 'Vietnam', 'Yemen')

south_america <- c('Argentina', 'Brazil', 'Colombia', 'Ecuador', 'Peru', 'Venezuela')
oceania <- c('Australia')
europe <- c('Belarus', 'Finland', 'France', 'Germany', 'Italy', 'Netherlands', 'Northern Ireland', 'Norway', 'Republic of Ireland', 'Russia', 'Turkey', 'UK', 'Ukraine', 'Wales, UK')
africa <- c('Benin', 'DR Congo', 'Egypt', 'Ethiopia', 'Kenya', 'Morocco', 'Mozambique', 'Nigeria', 'Sierra Leone', 'Somalia', 'Somaliland', 'South Africa', 'Tanzania', 'Uganda', 'Zambia', 'Zimbabwe')
north_america <- c('El Salvador', 'Jamaica', 'Mexico', 'US')

# add new column for continent
women <- women %>%
    mutate(continent = NA) 

# add continents to women dataframe
women$continent <- ifelse(women$country %in% asia, 'Asia', women$continent)
women$continent <- ifelse(women$country %in% south_america, 'South America', women$continent)
women$continent <- ifelse(women$country %in% oceania, 'Oceania', women$continent)
women$continent <- ifelse(women$country %in% europe, 'Europe', women$continent)
women$continent <- ifelse(women$country %in% africa, 'Africa', women$continent)
women$continent <- ifelse(women$country %in% north_america, 'North America', women$continent)

women
```

## Data Wrangling

The key to using the `sunburstR` package with this specific dataset is the wrangling that happens to filter by continents we created above. We'll also want to get rid of dashes with `mutate_at` as dashes are structurally needed to render the sunburst plots.

Below, I've filtered the `women` data frame into Africa and Asia (the same could be done for North and South America and Europe as well). 

The **two most important** operations here are the creation of the `path` and `V2` columns that will later be parameters for rendering the sunburst plots. 

```{r wrangling, message=FALSE}

# Filter for Africa
africa_name <- women %>%
    select(continent, category, role, name) %>%
    # remove dash within dplyr pipe
    mutate_at(vars(3, 4), funs(gsub("-", "", .))) %>% 
    filter(continent=='Africa') %>%
    mutate(
        path = paste(continent, category, role, name, sep = "-")
    ) %>%
    slice(2:100) %>%
    mutate(
        V2 = 1
    )

# Filter for Asia
asia_name <- women %>%
    select(continent, category, role, name) %>%
    # remove dash within dplyr pipe
    mutate_at(vars(3, 4), funs(gsub("-", "", .))) %>%
    filter(continent=='Asia') %>%
    mutate(
        path = paste(continent, category, role, name, sep = "-")
    ) %>%
    slice(2:100) %>%
    mutate(
        V2 = 1
    )

```

## Sunburst: Africa

Ultimately, I found the information best presented by continent as the *base* of the sunburst plot, followed by category, specific roles and the names of each of the 100 women honored by the BBC. 

Moreover, by presenting the data by continent, you can focus on just five specific color as you decide on a palette. 

I wouldn't recommend trying to pick a color for each role or name; it becomes too unweildy. Just pick five colors for the two inner most rings of the sunburst plot and it'll shuffle the rest of the colors.

```{r africa}
# Africa
sunburst(data = data.frame(xtabs(V2~path, africa_name)), legend = FALSE,
         colors = c("D99527", "6F7239", "CE4B3C", "C8AC70", "018A9D"))

```

## Sunburst: Asia

```{r asia}
# Asia
sunburst(data = data.frame(xtabs(V2~path, asia_name)), legend = FALSE,
         colors = c("#e6e0ae", "#dfbc5e", "#ee6146", "#d73c37", "#b51f09"))
```

And that's it for visualizing the BBC's top 100 influential women in 2020 with the `sunburstR` package. 

For more content on data science, visualization, in R and Python, [find me on Twitter](https://twitter.com/paulapivat). 